Co-moving stars in \textsl{Gaia DR1}

### 2016-11-10 paper/ms.tex
Modify figure 1 and 3
- new macro for consistant dvtan symbol
- combine figure 1 and 2 and revmoe 2
- change "candidate pair list" to "initial sample of pairs"
- change "network" to "connected component"

% define macros for math
\newcommand{\given}{\,|\,}
\newcommand{\normal}{{\mathcal{N}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transp}[1]{{#1}^{\!\mathsf{T}}}
\newcommand{\inv}[1]{{#1}^{-1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\vperp}{\bs{v}^\perp}
\newcommand{\propm}{\bs{\mu}}
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\bs{#1}}
\newcommand{\kms}{\ensuremath{\rm km~s^{-1}}}
\newcommand{\msun}{{\rm M}_\odot}
\newcommand{\data}{\mathrm{data}}
\newcommand{\snr}{[S/N]_\varpi}
\newcommand{\eye}{\mathbb{I}}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
creates an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here we seize this opportunity.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with surprisingly similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in the OB associations.
However, there still is a huge number of very wide ($>3$~pc) separation pairs of
stars that are co-moving.
We discuss their nature, and the prospects for testing stellar models and chemical
abundance measurements using our catalog.

@@

We construct a sample of candidate pair list as follows.
We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
Then, for each star, we search a potential co-moving partner star with a
difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
using RA, DEC, a point-estimate of the distance, and the proper motions.
Throughout this \documentname\, we use the following estimator for
the point-estimate of the distance applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  d = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
where $\varpi$ is the parallax in mas.
The difference in tangential velocity between two stars is, then,
\begin{equation}
  |\Delta \vec{v}_{\rm t}| = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

Figure~\ref{fig:dv-sep} shows $|\Delta v_t|$ against the physical separation
for the resulting 271,232 unique candidate pairs.
A few key observations can be made:

@@

  \item  However, it is not only (wide) binaries which are gravitationally
  bound that are expected to appear as co-moving.
  OB associations, moving groups, and open clusters are also of consideration.
  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
  As the pair separation increases, the nature of co-moving pairs
  will change from binaries to those related to these larger objects,
  which generally subtend a larger angle in sky.
  Since the proper motions of two stars with same 3D velocity
  are essentially projections of this velocity onto sphere at
  two different viewing angles,
  the larger the difference in viewing angles is, the larger the difference in tangential 
  velocities will be.
  Due to this projection effect, a population of genuine co-moving pairs
  will extend to larger $\Delta v_t$ at larger separation.
  This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
  in lower right corner that gets thinner as $\Delta v_t$ increases.

  \item Finally, there is a population of “random” pairs of field stars
  which are not actually co-moving, but still has $\Delta v_t < 10$~\kms\
  by chance.
  As $\Delta v_t$ increases, this population will dominate.
  However, it is clear from this Figure that there is an overlap between
  genuine co-moving pairs and “random” pairs.
\end{itemize}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
  \end{center}
  \caption{%
    Difference in tangential velocity vs. physical separation for candidate
    pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
    \label{fig:dv-sep}}
\end{figure*}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/a_vorb.pdf}
  \end{center}
  \caption{%
    Binary orbital velocity as a function of semi-major axis for $2~M_\odot$
    binary.
    \label{fig:vorbit}}
\end{figure*}

For wide binaries, the assumption of same 3D velocity can break down
for high precision proper motion measurements because of the orbital
velocity (Figure~\ref{fig:vorbit}).
To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.

>>>>

% define macros for math
\newcommand{\given}{\,|\,}
\newcommand{\normal}{{\mathcal{N}}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transp}[1]{{#1}^{\!\mathsf{T}}}
\newcommand{\inv}[1]{{#1}^{-1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\vperp}{\bs{v}^\perp}
\newcommand{\propm}{\bs{\mu}}
\newcommand{\mat}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\bs{#1}}
\newcommand{\kms}{\ensuremath{\rm km~s^{-1}}}
\newcommand{\msun}{{\rm M}_\odot}
\newcommand{\data}{\mathrm{data}}
\newcommand{\snr}{[S/N]_\varpi}
\newcommand{\eye}{\mathbb{I}}
\newcommand{\absdvtan}{\ensuremath{|\Delta\vec v_\mathrm{t}|}}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
creates an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here we seize this opportunity.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with surprisingly similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in OB associations.
However, there still is a huge number of very wide ($>3$~pc) separation pairs of
stars that are co-moving.
We discuss their nature, and the prospects for testing stellar models and chemical
abundance measurements using our catalog.

% change to probabilistic statement "co-moving"
% we present a table of co-moving pairs
% add dynamical inference 
% the last line = the most important thing

@@

We construct the initial sample of pairs as follows.
We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
Then, for each star, we search a potential co-moving partner star with a
difference in tangential velocity $\absdvtan < 10$~\kms\ within 10 pc
using RA, DEC, a point-estimate of the distance, and the proper motions.
Throughout this \documentname\, we use the following estimator for
the point-estimate of the distance applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  d = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
where $\varpi$ is the parallax in mas.
The difference in tangential velocity between two stars is, then,
\begin{equation}
  \absdvtan = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

Figure~\ref{fig:dv-sep} shows \absdvtan\ against the physical separation
for the resulting 271,232 unique pairs in the initial sample.
A few key observations can be made:

@@

  \item  However, it is not only (wide) binaries which are gravitationally
  bound that are expected to appear as co-moving.
  OB associations, moving groups, and open clusters are also of consideration.
  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
  As the pair separation increases, the nature of co-moving pairs
  will change from binaries to those related to these larger objects,
  which generally subtend a larger angle in sky.
  Since the proper motions of two stars with same 3D velocity
  are essentially projections of this velocity onto sphere at
  two different viewing angles,
  the larger the difference in viewing angles is, the larger the difference in tangential 
  velocities will be.
  Due to this projection effect, a population of genuine co-moving pairs
  will extend to larger \absdvtan\ at larger separation.
  This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
  in lower right corner that gets thinner as \absdvtan\ increases.

  \item Finally, there is a population of “random” pairs of field stars
  which are not actually co-moving, but still has $\absdvtan 10$~\kms\
  by chance.
  As \absdvtan\ increases, this population will dominate.
  However, it is clear from this Figure that there is an overlap between
  genuine co-moving pairs and “random” pairs.
\end{itemize}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
  \end{center}
  \caption{%
    Initial sample of pairs on the
    difference in tangential velocity vs. physical separation plane.
    For the initial sample, we look for stars
    with separation $< 10$~pc and $\absdvtan < 10$~\kms\ for
    every star in \tgas.
    The blue solid line shows the magnitude of the 3D
    orbital velocity as a function of semi-major axis for a $2~M_\odot$ binary system.
    Note the stream of points that starts at small separation ($\lesssim 0.01$~pc),
    small $\absdvtan$ ($\lesssim 2$~\kms),
    but climbs to larger $\absdvtan$ at $|\Delta \vec{x}|\gtrsim 1$~pc, which
    eventually merges with random pairs of field stars dominating in the upper right corner
    (see \S~\ref{sec:data} for details).
    \label{fig:dv-sep}}
\end{figure*}

@@

For wide binaries, the assumption of same 3D velocity can break down
for high precision proper motion measurements because of the orbital
velocity (blue solid line in Figure~\ref{fig:dv-sep}).
To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
